### 2017-01-19 robust-software.tex
Added example set to rule 9, some other tweaks to grammar and flow.

Being unable to
replicate results is so common that one publication refers to it as ``a rite of
passage''~\cite{baker2016}. 
The root cause of this problem is that most research software
is essentially a prototype, and therefore is not 
\emph{robust}. The lack of
robustness in published, distributed software
leads to duplicated efforts with little practical benefit,
which slows the pace of research~\cite{prabhu2011,lawlor2015}. 
Not all software needs to be robust enough to run for other 
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools to perform the same task:
for example,
in 2016 the Bioinformatics Links Directory included 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

@@

Our focus in this short guide is on the software that is already built,
not on the process used to build it,
but we believe that version control is essential to sustainable software development
\cite{wilson2014,wilson2016}.
In particular,
developers will struggle to understand what they have actually built,
what it actually does,
and what they have actually released
without some mechanical way to keep track of changes.
They should therefore
\ruleminor{put everything created manually into version control as soon as it is created},
including programs, original field observations, and the source files for papers.
Files that can be regenerated at need,
such as the binaries for compiled programs or intermediate files generated during data analysis,
should not be versioned;
instead,
it is often more sensible to use an archiving system for them,
and store the metadata describing their contents in version control instead~\cite{noble2009}.

@@

\rulemajor{Rule 5: Reuse software (within reason).}

@@

In order to be useful, \ruleminor{make the test script easy to find and run}. 
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a 
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a much better solution than documenting
test cases and requiring people to copy and paste them.

>>>>

Being unable to
replicate results is so common that one publication refers to it as ``a rite of
passage''~\cite{baker2016}.
The root cause of this problem is that most research software
is essentially a prototype, and therefore is not
\emph{robust}. The lack of
robustness in published, distributed software
leads to duplicated efforts with little practical benefit,
which slows the pace of research~\cite{prabhu2011,lawlor2015}.
Not all software needs to be robust enough to run for other people,
but software that is published and distributed to the community should be.
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools that perform similar tasks:
for example,
in 2016 the Bioinformatics Links Directory included 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

@@

Version control is essential to sustainable software development
\cite{wilson2014,wilson2016}.
In particular,
developers will struggle to understand what they have actually built,
what it actually does,
and what they have actually released
without some mechanical way to keep track of changes.
They should therefore
\ruleminor{put everything created manually into version control as soon as it is created},
including programs, original field observations, and the source files for papers.
Files that can be regenerated at need,
such as the binaries for compiled programs or intermediate files generated during data analysis,
should not be versioned;
instead,
it is often more sensible to use an archiving system for them,
and store the metadata describing their contents in version control instead~\cite{noble2009}.

If you are new to version control,
it is simplest to treat it as ``a better \withurl{Dropbox}{http://dropbox.com}''
(or, if you are of a certain age, a better FTP),
and to use it simply to synchronize files between multiple developers and machines~\cite{blischak2016}.
Once you are comfortable working that way,
you should \ruleminor{use a feature branch workflow}:
designate one parallel copy (or ``branch'') of the repository as the master,
and create a new branch from it each time you want to fix a bug or add a new feature.
This allows work on independent changes to proceed in isolation;
once the work has been completed and tested,
it can be merged into the master branch for release.

@@

\rulemajor{Rule 5: Reuse software (within reason)}

@@

In order to be useful, \ruleminor{make the test script easy to find and run}.
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a 
good solution that lets new users build their analysis from a working script.
For example, with its distribution, 
HISAT2 includes a full set of very small files,
and a 'Getting Started with HISAT2' section in its manual 
that leads you through the entire data lifecycle~\cite{pertea2016}.
